---
title: 'UCI data : Online Retail'
author: "류성균"
date: '2021 1 6 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-   reference

    -   UCI repository : [Online Retail](http://archive.ics.uci.edu/ml/datasets/online+retail)

    -   Datacamp : [**Market Basket Analysis using R**](https://www.datacamp.com/community/tutorials/market-basket-analysis-r)

    -   [토픽모형을 이용한 빅데이터 기반 마이크로 세그멘테이션 방법론 연구(김용대, 정구환, 2016)](http://kostat.go.kr/file_total/ex-1-02.pdf)

```{r}
#install and load package arules
#install.packages("arules")
# library(arules)
#install and load arulesViz
#install.packages("arulesViz")
# library(arulesViz)
#install and load tidyverse
#install.packages("tidyverse")
library(tidyverse)
#install and load readxml
#install.packages("readxml")
library(readxl)
#install and load knitr
#install.packages("knitr")
library(knitr)
#load ggplot2 as it comes in tidyverse
library(ggplot2)
#install and load lubridate
#install.packages("lubridate")
library(lubridate)
#install and load plyr
#install.packages("plyr")
library(plyr)
library(dplyr)
library(here)
library(lda)
```

## Data Description

#### **Dataset Description**

-   **Number of Rows:**541,909

-   **Number of Attributes:**08

##### Attribute Information

-   InvoiceNo: Invoice number. Nominal, a 6-digit integral number uniquely assigned to each transaction. If this code starts with letter 'c', it indicates a cancellation. +StockCode: Product (item) code. Nominal, a 5-digit integral number uniquely assigned to each distinct product.

-   Description: Product (item) name. Nominal.

-   Quantity: The quantities of each product (item) per transaction. Numeric.

-   InvoiceDate: Invoice Date and time. Numeric, the day and time when each transaction was generated. Example from dataset: 12/1/2010 8:26

-   UnitPrice: Unit price. Numeric, Product price per unit in sterling.

-   CustomerID: Customer number. Nominal, a 5-digit integral number uniquely assigned to each customer.

-   Country: Country name. Nominal, the name of the country where each customer resides.

## Data Pre-Processing

```{r}
# read excel into R dataframe

retail <- read_excel(here('data/online_retail/Online Retail.xlsx'))

# complete.cases(data) will return a logical vector indicating which rows have no missing value Then use the vector to get only rows that are complete using retail[,].
retail <- retail[complete.cases(retail),]

#mutate function is from dplyr package. It is used to edit or add new columns to dataframe. Here Description column is being converted to factor column. as.factor converts column to factor column. %>% is an operator with which you may pipe values to another function or expression
# retail <- retail %>% mutate(Description = as.factor(Description))
```

```{r}
# retail <- retail %>% mutate(Country = as.factor(Country))
```

```{r}
# Converts character data to date. Store InvoiceDate as date in new variable
retail$Date <- as.Date(retail$InvoiceDate)
# Extract time from InvoiceDate and store in another variable
TransTime <- format(retail$InvoiceDate, "%H:%M:%S")
# Convert and edit InvoiceNo into numeric
retail$InvoiceNo <- as.numeric(as.character(retail$InvoiceNo))
```

```{r}
#Bind new columns TransTime and InvoiceNo into dataframe retail
retail <- cbind(retail, TransTime)
```

```{r}
retail %>% glimpse()
```

```{r}
# 데이터는 약 일년치 데이터(2010-12-01~2011-12-09)
retail %>% 
  select(Date) %>% arrange(Date) %>% head()
retail %>% 
  select(Date) %>% arrange(Date) %>% tail()
```

```{r}
### 나중에 고민
# duplicated(retail) %>% table()
# 
# retail[duplicated(retail), ] %>% 
#   arrange(InvoiceNo, StockCode, Description)
```

```{r}
shop <- retail %>% 
  select(CustomerID, Description) %>% 
  dplyr::group_by(CustomerID, Description) %>% 
  dplyr::summarize(cnt = n()) %>% ungroup()
```

```{r}
### Description당 고유번호 확인 
Description <- 
  shop %>% 
    dplyr::select(Description) %>% 
    dplyr::arrange(Description) %>% 
    unique() 

Description %<>% 
  bind_cols(1:dim(Description)[1])

colnames(Description) <- c("Description", "No")

Description %>% tail(100)

### shop의 Description을 고유번호로 대체
shop <- shop %>% 
  left_join(Description, by="Description") %>% 
  select(CustomerID, No, cnt)
```

```{r}
write_csv(shop, here("data/online_retail/shop.csv"))

write.table(Description$Description, here("data/online_retail/Description.txt"))
```

## 토픽모델링

-   코드 출처 : [토픽모형을 이용한 빅데이터 기반 마이크로 세그멘테이션 방법론 연구(김용대, 정구환, 2016)](http://kostat.go.kr/file_total/ex-1-02.pdf)

```{r}
shop = read_csv(here("data/online_retail/shop.csv"))
mj = table(shop[,1]) # CustomerID count
outfile = file(here("data/online_retail/shop.dat"))
line = c(); k = 1
for(j in 1:length(mj)){ # customerID길이의 반복문
   line=paste0(line, mj[j])
   for(i in k:(k+mj[j]-1)){
     line=paste0(line, " ", shop[i, 2]-1, ":", shop[i, 3])
   }
   line=paste0(line, "\n")
   k=k+mj[j]
}
writeLines(line, outfile)
close(outfile)
```

```{r}
library(lda)
shop2 <- read.documents(here("data/online_retail/shop.dat"))
item_name <- read.vocab(here("data/online_retail/Description.txt"))

K <- 11; alpha <- 1.0; beta <- 1.0; iter = 500
lda <- lda.collapsed.gibbs.sampler(shop2,
                                   K, 
                                   item_name,
                                   iter,
                                   alpha,
                                   beta,
                                   compute.log.likelihood =  T)

```

#### 자료 분석결과

- $\theta$와 $\phi$ 추정결과

```{r}

plot = data.frame(iteration = c(1:iter, 1:iter),
                  loglike = c(lda$log.likelihoods[1,],
                              lda$log.likelihoods[2,]),
                  type = c(rep(1,iter), rep(2,iter)))

ggplot(plot, aes(x = iteration, y = loglike, group = type)) + 
  geom_line()
```
- $\theta$와 $\phi$를 추정하는 R코드
```{r}
n = 4372 # 고객 수
W = dim(Description)[1] + 1
theta = matrix(0, nrow=n, ncol=K)

for (i in 1:n){
   theta[i,] = t(lda$document_sums[,i])/sum(lda$document_sums[,i])
} 
phi = matrix(0, nrow=K, ncol = W)

for (i in 1:K){
  phi[i, ] = lda$topics[i,]/sum(lda$topics[i,])
}

theta %>% summary()
# phi %>% summary()
```

##### 리프트 계산
```{r}
p = colSums(lda$topics) / sum(lda$topics)
lift = matrix(0, nrow = K, ncol = W)
colnames(lift) <- item_name

topic_name = c()

for (i in 1:K){
  lift[i, ] <- phi[i, ] / p
  sorted = sort(lift[i,], decreasing = T )[1:2]
  topic_name = c(topic_name,
                 paste(names(sorted), collapse = "."))
}
```

##### 토픽 그리기
```{r}
library(ggplot2)
library(reshape2)
theme_set(theme_bw())
colnames(phi) <- item_name; idx = c(1, 6, 7)
phi.df <- melt(cbind(data.frame(phi[idx,]),
                     topic = factor(idx)), variable.name = "item",
               id.vars = "topic")


qplot(item, value, fill = item, ylab = "value", data = phi.df,
      geom = "bar", stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") +
  coord_flip() + 
  facet_wrap(~ topic, ncol = length(idx))
```

